\chapter{Literature review}
\section{Chapter overview}

This chapter situates the proposed research within two long-standing
traditions of stock analysis. It begins by revisiting the conceptual
distinction between \emph{fundamental analysis} and \emph{technical
analysis}, drawing on \citeauthor{Graham2006}'s foundational separation
of investment from speculation \citep{Graham2006}, before mapping how
recent advances in agentic artificial intelligence (AI) have begun to
reshape both. The review is structured thematically rather than
paper-by-paper: it first establishes the problem domain and the gap
that fundamental analysis poses for automation, then critically examines
the technological building blocks of agentic systems, before turning to
existing work on (i)~agentic systems for fundamental analysis, which
is the focus of the proposed system, and (ii)~agentic approaches to
sentiment analysis, which dominate the technical-analysis side of the
literature. The chapter closes with a critical view of evaluation
practice and a synthesis.

% \section{Concept map}
% Insert a conceptual map figure and provide a brief explanation of concept relationships.

% \begin{figure}[H]
%     \centering
%     \fbox{\rule{0pt}{2in}\rule{0.9\linewidth}{0pt}}
%     \caption{Concept map of the research domain (placeholder)}
%     \label{fig:concept-map}
% \end{figure}

\section{Problem domain review}

\subsection{Two Traditions: Fundamental Analysis and Price Prediction}
\label{sec:twotraditions}

Modern stock analysis still rests on the dichotomy that
Benjamin Graham drew more than seventy years ago. In
\emph{The Intelligent Investor}, \citet{Graham2006} argues that an
investment operation is one which, on thorough analysis, promises
safety of principal and an adequate return, while operations not
meeting these requirements are speculative. Fundamental analysis,
in this tradition, is the disciplined assessment of a business based on
its earnings power, balance-sheet strength, competitive position
and future profitability and is deliberately price-agnostic;
the share price enters the picture only at the end, when the analyst
compares intrinsic value to the quoted price to decide whether a
margin of safety exists \citep{Graham2006}. Technical analysis takes
the opposite stance: it treats the price series itself as the primary
object of study and tries to forecast future prices from past prices,
volumes and, increasingly, the mood of market participants
\citep{murphy1999}. \citet{Graham2006} is unambiguous that activities
of this second kind, however sophisticated, fall on the speculative
side of the line.

This distinction matters for the design of any AI-driven analyst,
because the two traditions ask categorically different questions.
A fundamental engine is asked, in effect, what the company is worth;
a technical engine is asked where the price is heading next. Recent
agentic systems often blur this line by bundling fundamental,
sentiment and technical signals into a single recommendation
\citep{mattaparty2025,setyowidodo2025}, but the underlying epistemic
difference persists, and it is the choice of which question to
answer that determines what data, what reasoning and what evaluation
are appropriate.

\subsection{Why Fundamental Analysis Resists Automation}
\label{sec:whyresist}

Fundamental analysis is a deceptively difficult task to automate.
The raw inputs like annual reports, 10-K and 10-Q filings,
earnings-call transcripts, management commentary, industry studies
and macroeconomic releases are heterogeneous, long, frequently
qualitative, and arrive at different cadences
\citep{Fatouros2025MarketSense}. Equity-research practice typically combines
a study of the business model, the financial statements and recent
conference calls into a single judgement \citep{dongare2024}, an
integration that is hard to encode in conventional rule-based or
purely statistical pipelines. Earlier machine-learning approaches
sidestepped this difficulty by reducing fundamentals to a small set
of accounting ratios and treating the problem as a classification
of price direction \citep{bustos2020}, which arguably crosses
Graham's line into speculation by tying the analysis directly to
short-horizon price moves.

Even for large language models (LLMs), three obstacles recur. First,
context-window limits prevent the direct ingestion of long filings
such as 10-Ks or detailed macroeconomic reports
\citep{Fatouros2025MarketSense}. Second, LLMs are known to hallucinate
numerical values and to handle tables inconsistently, which is a
serious problem in financial fundamentals where precision matters
\citep{saravanan2026}. Third, monolithic LLMs conflate causation
with correlation when reasoning over financial time series,
undermining the auditability that fundamental decisions require
\citep{saravanan2026}. These limitations have pushed researchers
towards agentic architectures that decompose fundamental analysis
into specialised sub-tasks.

\subsection{Frameworks Shaping the Design Space}
\label{sec:frameworks}

Several frameworks now shape how agentic systems are built in this
space. Multi-agent orchestration libraries such as AutoGen and
CrewAI provide patterns for role-based collaboration, tool use and
conversational delegation \citep{saravanan2026,setyowidodo2025}.
Retrieval-Augmented Generation (RAG) has become the standard
mechanism for grounding LLM reasoning in long financial documents
\citep{han2024}. On the empirical side, financial-domain language
models such as FinBERT \citep{huang2023finbert} and finance-specific
large models such as BloombergGPT and FinGPT have established
reusable text-encoding components that downstream agentic systems
routinely call upon \citep{saravanan2026}. Surveys of LLM
applications in finance offer a useful map of the territory but are
necessarily broad rather than prescriptive \citep{liu2024}.
Industry-facing reviews further confirm that agentic AI is moving
from an experimental capability to a deployment-stage technology in
financial services, with autonomy, adaptability and coordination
identified as its three defining characteristics \citep{bhat2025}.

\subsection{Positioning of the Proposed System}
\label{sec:positioning}

Against this backdrop, the proposed system focuses deliberately on
the fundamental side of Graham's divide. It adopts an agentic
architecture in which specialised agents read filings, financial
statements, earnings calls and macroeconomic context, and synthesise
these into a view of business quality and future profitability without taking a stand on near-term price. Sentiment, which is
treated extensively in the literature as a driver of price prediction
\citep{bollen2011,tetlock2007}, is therefore reviewed in
Section~\ref{sec:sentimentagents} as part of the technical-analysis
tradition, not as a core component of the proposed fundamental
engine.

% figures showing proposed architecture

\section{Technological Review}
\label{sec:techreview}

\subsection{Large Language Models as Reasoning Engines}
\label{sec:llms}

LLMs have become the de facto reasoning substrate for agentic systems
in finance. Their appeal lies in three capabilities: ingesting long,
mixed-modal text; generating natural-language rationales; and
following structured prompts. \citet{Papasotiriou2024} show that
GPT-4, when prompted in a Chain-of-Thought style and given the same
inputs as a human analyst, can produce stock ratings that beat
consensus analyst ratings on forward-return mean absolute error.
Critically, they find that the most accurate variant is the one fed
company fundamentals, with news and sentiment offering only marginal
short-term gains \citep{Papasotiriou2024}. This empirical result is
consistent with Graham's claim that fundamentals carry the
long-horizon signal, while sentiment is a short-horizon overlay
\citep{Graham2006}.

Backbone choice nevertheless matters. \citet{yu2025finmem} compare
GPT-3.5, GPT-4, GPT-4-Turbo and Llama-2 inside the same agent
framework and report that GPT-4-class models consistently dominate on
cumulative return and Sharpe ratio. Such backbone sensitivity is a
reminder that LLM-based fundamental analysis cannot be evaluated in
isolation from the underlying model, and that reproducibility
requires explicit reporting of model versions, prompts and
temperature settings.

\subsection{Multi-Agent Orchestration Patterns}
\label{sec:agentpatterns}

Once an LLM is treated as a tool rather than a black-box solver, the
question becomes how to coordinate several of them. \citet{han2024}
formalise this as a question of agent group size and collaboration
structure, and find empirically that single agents outperform
multi-agent debates for fundamental analysis specifically, while
ensemble structures dominate for the higher-level "buy or not"
decision, achieving 66.7~per~cent binary accuracy and a 2.35~per~cent
target-price error on Dow Jones 10-K filings. Building on these
results, \citet{chen2025dda} propose a directed dual-agent design
in which a subordinate agent specialises in fundamentals, sentiment
or risk and a leader agent synthesises the output, reporting
68.3~per~cent buy/not-buy accuracy and a 2.18~per~cent price error
under a strong-leadership configuration with evidence that lean,
vertically structured collaborations can outperform larger horizontal
teams when coordination overhead is high.

At the other end of the design spectrum, the GWise framework treats
agents as nodes in a directed acyclic graph, with an LLM orchestrator
selecting a subgraph and softmax-weighting agent outputs into a final
trade signal \citep{rezaei2025gwise}. \citet{Cheng2025} similarly
deploy specialised analyst, strategist, trader and risk-management
agents within an AutoGen/CrewAI pipeline, reporting a Sharpe ratio
of 2.38 against a buy-and-hold baseline of 1.23 on AAPL. These
designs share a common premise that the structure of agent
communication is itself a design variable even though they answer
it differently.

\subsection{Retrieval, Memory and Domain Grounding}
\label{sec:retrievalmemory}

Two technical mechanisms recur in agentic finance systems:
retrieval-augmented generation and explicit memory. RAG provides a
way to ground LLM reasoning in documents that exceed the model's
context window; \citet{han2024} encapsulate retrieval as a tool
function that agents call autonomously, while \citet{Fatouros2025MarketSense}
extend basic RAG with semantic chunking and Hypothetical Dense
Embeddings (HyDE) to extract macroeconomic context from institutional
reports, reporting precision of 1.00 and recall of 0.91 at top-7
chunks. Memory mechanisms address a different limitation: the
inability of stateless LLMs to retain time-varying signals across
trading days. The FINMEM agent introduces a layered memory module that is shallow, intermediate and deep with distinct decay rates for
daily news, quarterly 10-Q insights and annual 10-K insights,
retrieved via similarity, recency and importance scores
\citep{yu2025finmem}. The architecture is interesting because it
consciously mirrors the cognitive structure of a human trader, and
ablation studies show that both the memory module and the dynamic
risk-profile setting contribute non-trivially to performance
\citep{yu2025finmem}.

\section{Existing Work}
\label{sec:thematicreview}

\subsection{Agentic AI for Fundamental Analysis}
\label{sec:fundamentalagents}

The most directly relevant strand of literature places fundamental
analysis at the centre of an agentic pipeline. MarketSenseAI~2.0 is
the most ambitious example: a five-agent system (News, Fundamentals,
Dynamics, Macroeconomic, Signal) in which the Fundamentals agent
itself orchestrates a three-layer chain over SEC filings,
earnings-call transcripts and quarterly numerical data
\citep{Fatouros2025MarketSense}. On the S\&P 100 over 2023--2024 this design
produces a cumulative return of 125.9~per~cent against the index
return of 73.5~per~cent, with a 16~per~cent higher Sortino ratio and
an explicit alpha of 8.0--18.9~per~cent on the broader S\&P 500
\citep{Fatouros2025MarketSense}. The result is notable not because of the
headline return backtests of this length are easily flattered by
survivorship bias but because the authors decompose returns
using Carhart and Fama-French factor models and still find
unexplained alpha attributable to the framework's qualitative
reasoning.

Other systems take a chatbot-first stance on fundamentals.
\citet{dongare2024} build an LLM chatbot that ingests
profit-and-loss statements, balance sheets, cash-flow reports and
conference-call transcripts as PDFs, and answers fact-based,
behavioural and advice-seeking questions with differing accuracy
across question types. The system explicitly refuses speculative
investment advice, which is in keeping with Graham's normative stance
but limits its utility as a stand-alone analyst. FinQAPT pushes the
same idea further with a three-module pipeline (FinPrimary,
FinContext, FinReader) and a Dynamic N-shot Prompting method that
achieves state-of-the-art 80.6~per~cent on the FinQA dataset at
module level \citep{singh2024finqapt}. At the other end of the
integration spectrum, \citet{saravanan2026} propose MAC, a
neuro-symbolic three-agent system in which an Analyst harmonises
noisy SME transactions to a financial ontology, a Quant performs
causal inference using vector autoregression, Granger causality and
the PC algorithm, and a Partner generates a strategic narrative
grounded only in verified causal facts. The deliberate confinement
of generative behaviour to the explanation stage is the most explicit
attempt yet in this literature to address Graham's concern about
disciplined, auditable analysis.

A closely related cluster of work uses agents to reason over annual
filings directly. \citet{han2024} demonstrate that a single
fundamental agent reading a 10-K outperforms multi-agent debates for
fundamental sub-tasks, and that ensemble strategies achieve the best
target-price prediction error of 2.35~per~cent.
\citet{mattaparty2025} combine personality-style agents inspired by
Graham and Buffett with valuation, technical and sentiment agents in
a unified pipeline; on AAPL the system reports an aggregate
confidence score of 70.5~per~cent and a margin-of-safety calculation
of $-88.51$~per~cent for the Graham agent, suggesting overvaluation.
Although the architecture is wide-ranging, the use of a Graham-styled
agent is symptomatic of how the field is reaching back to classical
investment thought to discipline LLM behaviour. The Equity Research
Chatbot of \citet{dongare2024} and the broader survey of agentic AI
in finance by \citet{bhat2025} confirm that fundamental analysis is
widely seen as the natural test bed for serious agentic deployments.

\subsection{Agentic Approaches to Sentiment Analysis}
\label{sec:sentimentagents}

Sentiment analysis sits squarely on the technical-analysis side of
Graham's divide because its purpose is to predict short-horizon
price movements from the mood of market participants. The literature
here is much larger than for fundamentals, and four sub-themes are
worth distinguishing.

\subsubsection{News-Driven Sentiment Agents}
\label{sec:newssentiment}

The earliest and still most common approach is to score financial
news for tone and use this as a feature for a downstream model. The
classical evidence that news content drives short-term returns is
\citet{tetlock2007}. FinBERT \citep{huang2023finbert} operationalised
this insight as a pre-trained encoder that has since become the
default sentiment backbone for agentic systems. In recent agentic
settings, sentiment is rarely a stand-alone module: it is one
specialised agent among several. The News Agent in MarketSenseAI~2.0,
for instance, condenses daily news into a progressive narrative that
the downstream Signal Agent uses for short-term momentum
\citep{Fatouros2025MarketSense}, and the Sentiment Analyst in the
\citet{Cheng2025} framework generates tone scores that are fused with
fundamental and technical signals before a trading decision is taken.
Critically, when \citet{Papasotiriou2024} test sentiment in
isolation, they find only a modest short-term predictive benefit and
no significant gain over fundamentals at twelve and eighteen month
horizons, an empirical reminder that sentiment's natural domain
is short-horizon price prediction, not long-horizon valuation.

\subsubsection{Social-Media and Retail-Investor Sentiment Agents}
\label{sec:socialsentiment}

A second sub-theme draws on the long-running line of work that began
with \citet{bollen2011}, who showed that aggregate Twitter mood
predicts the Dow Jones Industrial Average with a lag of three to four
days. Modern agentic systems institutionalise this signal. The Social
Media Analyst in the GWise framework parses Twitter, Reddit and
StockTwits for retail tone \citep{rezaei2025gwise}, and the Sentiment
Analyst Agent in \citet{mattaparty2025} extracts public sentiment
from financial news and social media using FinBERT. These systems
treat retail sentiment as a fast-moving, noise-prone but informative
input to the trading decision rather than to the assessment of
business value. The agentic angle adds little to the underlying
signal but improves how it is aggregated and reconciled with other
inputs.

\subsubsection{Earnings-Call and Management-Tone Sentiment Agents}
\label{sec:earningssentiment}

A more recent and arguably more interesting strand uses agents to
read earnings-call transcripts for managerial tone. The Earnings Call
Analyst in the GWise framework applies natural language processing
to chief executive and chief financial officer language to detect
confidence or hesitation \citep{rezaei2025gwise}.
\citet{bharadwaj2025} take this further by combining FinBERT-derived
sentiment from quarterly reports, newsletters, earnings calls and
press releases with a reinforcement-learning trader that uses
Proximal Policy Optimisation, Deep Q-Network and Advantage
Actor-Critic algorithms. They report that the inclusion of sentiment
from authoritative financial sources improves directional accuracy,
echoing earlier findings on the temporal evolution of earnings-call
sentiment \citep{bharadwaj2025}. There is an interesting tension
here: although earnings calls are clearly a fundamental source of
information, sentiment-style processing of them collapses managerial
commentary into a positive/negative score, discarding precisely the
qualitative content that fundamental analysis would seek to
interpret.

\subsubsection{Heterogeneous Multi-LLM Sentiment Frameworks}
\label{sec:hadsentiment}

The fourth sub-theme is methodologically distinctive.
\citet{xing2025had} argues, drawing on Minsky's theory of mind and
emotions, that a single LLM cannot adequately capture the layered
emotional content of financial text. The Heterogeneous Agent
Discussion (HAD) framework instantiates several specialised LLM
agents, each guided by a different linguistic or financial
heuristic and uses a summative agent to reason over their
aggregated discussion \citep{xing2025had}. Across six financial
sentiment datasets the framework yields better accuracy than
alternative multi-LLM settings, especially when the discussion
content is substantial. The contribution is not a new sentiment
lexicon but a design pattern: agentic disagreement, properly
summarised, can be used as a feature rather than a defect. This
finding interacts with \citeauthor{han2024}'s \citeyearpar{han2024}
result that multi-agent debate is sometimes counterproductive for
simple sub-tasks; together they suggest that the value of
multi-agent architectures rises with task complexity, which sentiment
around investor psychology arguably has.

\subsection{Hybrid Systems and Cross-Paradigm Pipelines}
\label{sec:hybrid}

A growing number of systems explicitly mix fundamental, technical,
sentiment and macroeconomic inputs. \citet{setyowidodo2025} use the
CrewAI framework to orchestrate a fundamental agent (75~per~cent
reported accuracy on long-term value) alongside a technical agent
(50~per~cent, struggling with reversals) and a macroeconomic agent
for the Indonesian market. The authors candidly observe that the
technical agent is the weakest component, which is consistent with
the wider literature on the difficulty of price-direction prediction.
\citet{mashaly2025} integrate a multi-agent system over stock-market
news with classical time-series models such as SARIMAX, LSTM and
Time-MOE, using LLM-derived news ratings as exogenous variables in
the regression layer; the design effectively treats sentiment as a
covariate of price rather than as a substitute for fundamental
judgement. Hybrid designs of this kind respect Graham's distinction
operationally which is different signals enter the model in different
ways even when the language of the papers does not.

\section{Evaluation, Benchmarking and Critical Limitations}
\label{sec:evaluation}

A common feature of the empirical results above is generous
backtesting. Most LLM-based investing strategies are evaluated on
small numbers of well-known stocks (typically TSLA, AMZN, NFLX, MSFT)
over short windows of three to twelve months \citep{Li2026FINSABER}.
The recent FINSABER framework re-evaluates leading systems including
FinMem and FinAgent on twenty years of multi-source data with
explicit mitigation of survivorship, look-ahead and data-snooping
biases, and reports that the apparent superiority of these systems
disappears: neither FinMem nor FinAgent generates statistically
significant alpha ($p > 0.34$ in all settings tested), and both
underperform simple ARIMA and rule-based benchmarks on risk-adjusted
metrics across bull, sideways and bear regimes
\citep{Li2026FINSABER}. The implication is uncomfortable but
important: the field's headline numbers cannot be taken at face
value, and a serious agentic system aimed at fundamental analysis
must be evaluated over multi-year horizons with delisted stocks
included.

Datasets used across the work surveyed here are also worth noting.
Most studies rely on combinations of Yahoo Finance OHLCV data, Alpha
Vantage, SEC EDGAR filings, NewsData.io, Reddit r/WallStreetBets and
Bloomberg/Reuters news feeds
\citep{mattaparty2025,bharadwaj2025,yu2025finmem}. The diversity of
data supports rich agent design but complicates reproducibility,
since few systems release the underlying news corpora due to
licensing constraints. Evaluation metrics are increasingly
standardised like cumulative return, annualised return, Sharpe ratio,
Sortino ratio, maximum drawdown but task-level metrics for
fundamental analysis specifically (such as target-price mean absolute
error or buy/not-buy binary accuracy) remain study-specific
\citep{han2024}.

% =============================================================
\section{Chapter Summary}
\label{sec:summary}

Three takeaways shape the methodology that follows. First, Graham's
distinction between investment grounded in fundamentals and
speculation grounded in price prediction remains analytically useful
and is, if anything, sharpened by the new tools: agentic systems
make it visible whether a design is asking what a business is worth
or where its price will go next \citep{Graham2006}. Second, the
literature on agentic fundamental analysis like MarketSenseAI~2.0,
MAC, FinQAPT, and the agent-collaboration studies of Han \citep{han2024}
converges on a small set of design moves: specialise agents by data
source; ground generation in retrieval and, where possible, in
deterministic statistical methods; use memory and chain-of-thought
to handle the multi-timeliness of financial data; and reserve
generative behaviour for explanation rather than for numerical
computation \citep{Fatouros2025MarketSense,han2024,saravanan2026,singh2024finqapt}.
Third, sentiment analysis, however sophisticated its agentic
implementations, is empirically a short-horizon, price-side signal
and should be positioned accordingly
\citep{bollen2011,Papasotiriou2024,xing2025had}.

The proposed system, described in Chapter~3, draws on these lessons
to build a fundamental-analysis-first agentic architecture, in which
sentiment, if used at all, is a clearly delimited overlay rather than
the central reasoning engine.
