\chapter{Methodology}
\section{Chapter overview}

This chapter presents the methodological framework underpinning the design and development of the Agentic AI-based fundamental analysis system. It begins by outlining the research methodology through the lens of \citep{saunders2019research} Research Onion Model, justifying the philosophical, strategic, and procedural choices that guide the study. The chapter then describes the iterative development methodology adopted for system construction, followed by a discussion of the project management approach, which draws on Agile principles to support continuous refinement. The scope, schedule, resource requirements, and data requirements of the project are subsequently detailed, alongside a structured risk assessment with corresponding mitigation strategies. Together, these elements establish a coherent and rigorous methodological foundation for the research.

\section{Research methodology}

The following table provides an overview of the selected scientific methodology based on the
conventional Research ‘Onion’ Model, presented by \citep{saunders2019research}.

\begin{longtable}{|p{4.8cm}|p{10cm}|}
    \hline
    \textbf{Research Design Element} & \textbf{Description} \\
    \hline
    \endfirsthead

    \hline
    \textbf{Research Design Element} & \textbf{Description} \\
    \hline
    \endhead

    \hline
    \multicolumn{2}{|r|}{\textit{Continued on next page\ldots}} \\
    \hline
    \endfoot

    \hline
    \endlastfoot

    Research Philosophy & 
    This study adopts pragmatism as its research philosophy, as it 
    is primarily focused on addressing a real-world problem, enhancing 
    fundamental analysis through an Agentic AI-based approach, rather 
    than adhering to a single philosophical paradigm. Pragmatism enables 
    the integration of both quantitative methods and qualitative insights, 
    allowing the research to effectively handle diverse data types. \\
    \hline
    Research Approach & 
    An abductive research approach is employed, beginning with the 
    observation of limitations in conventional stock fundamental analysis 
    tools, and iteratively developing theoretical and practical insights 
    through the design and evaluation of an Agentic AI system. This 
    approach allows the research to move flexibly between data-driven 
    findings and theoretical reasoning throughout the development process. \\
    \hline
    Research Methodological Choice & 
    A mixed-methods design is adopted, combining quantitative analysis 
    of financial metrics with qualitative insights derived from news 
    articles, management commentary, and market narratives. This dual 
    approach enables a comprehensive evaluation of stock fundamentals 
    that neither method could achieve in isolation. \\
    \hline
    Research Strategy & 
    The research follows a design science strategy, centred on the 
    development and evaluation of an artefact, an Agentic AI-powered 
    fundamental analysis tool. The system is designed to autonomously 
    retrieve, process, and synthesise data, producing actionable 
    investment insights. Iterative prototyping and performance evaluation 
    against financial reports, broker research reports, and interviews 
    with domain experts form the core of the research strategy. \\
    \hline
    Research Time Horizon & 
    A cross-sectional time horizon is adopted, with the tool evaluated 
    against a defined snapshot of stock market data within a specific 
    period. This allows for the systematic assessment of the tool's 
    analytical capabilities without requiring longitudinal tracking, 
    though future work may extend to longitudinal evaluation. \\
    \hline
    Research Data Collection and Analysis & 
    Data is collected from multiple sources, including financial 
    statements from the Colombo Stock Exchange (CSE), news articles, 
    and analyst reports, which are analysed using natural language 
    processing techniques embedded within the Agentic AI pipeline. 
    The tool's outputs are evaluated based on accuracy, completeness, 
    and relevance of the generated fundamental analysis. \\
    \hline

    \caption{Research methodology}
    \label{tab:research-methodology}
\end{longtable}

\section{Development methodology}

This study adopts an iterative development methodology, where the proposed Agentic AI-based system is developed in incremental stages. Each iteration involves designing and implementing a prototype, followed by systematic evaluation of its performance in supporting fundamental analysis. The insights gained from each evaluation are used to refine and enhance the subsequent version of the system. This continuous cycle of development and improvement ensures that the solution evolves progressively, leading to a more effective and reliable framework for fundamental analysis. 

\section{Project management methodology}

Agile principles are adopted for project management to support iterative development and continuous improvement, enabling the system to evolve through successive refinements until it meets the desired project objectives.

\subsection{Project Scope}

This study focuses on fundamentally analysing publicly listed companies in the Colombo Stock Exchange (CSE). However, diversified holding companies are excluded due to the complexity arising from their multi-sector operations. The analysis is primarily based on annual reports, and quarterly reports are not considered. As a result, the findings may reflect a maximum time lag of up to one year. 

\subsection{Schedule}

\subsubsection{Gantt chart}

The schedule of the project is presented in the form of a Gantt chart, in APPENDIX B.1.

\subsubsection{Deliverables and dates}

\begin{table}[H]
    \centering
    \begin{tabular}{|l|l|}
        \hline
        Deliverable & Target date \\
        \hline
        Draft literature review  & Feb 28th, 2026  \\
        \hline
        Draft project proposal submission  & March 14th, 2026  \\
        \hline
        Project proposal presentation  & May 7th, 2026  \\
        \hline
        Project proposal and interim demonstration  & May 14th, 2026  \\
        \hline
        Software implementation completion & June 30th, 2026  \\
        \hline
        Thesis submission  & July 30th, 2026 \\
        \hline
    \end{tabular}
    \caption{Deliverables and dates}
    \label{tab:project-schedule}
\end{table}

\subsection{Resources}

\subsubsection{Software resources}


\begin{itemize}
    \item \textbf{Agentic AI Framework} (LangChain / LangGraph) --- used 
    for orchestrating multi-step AI agent pipelines for autonomous 
    retrieval, reasoning, and synthesis of financial data.

    \item \textbf{Large Language Model} (OpenAI GPT-5.5 / Claude 4.6 Sonnet / AWS Textract) 
    --- used for extracting qualitative insights, summarising financial 
    narratives, and interpreting unstructured financial text.

    \item \textbf{Vector Database} (ChromaDB / Pinecone) --- used for 
    storing and retrieving embedded financial documents for 
    Retrieval-Augmented Generation (RAG) within the agent pipeline.

    \item \textbf{Backend Development} (Python, FastAPI) --- used for 
    building the core application logic, API endpoints, and agent 
    orchestration layer.

    \item \textbf{NLP and Embeddings} (HuggingFace Transformers, OpenAI 
    Embeddings) --- used for generating text embeddings for semantic 
    search and qualitative document retrieval.

    \item \textbf{Frontend / Interface} (React) --- used for 
    developing the user interface for interacting with the 
    fundamental analysis tool.

    \item \textbf{Version Control} (Git, GitHub) --- used for source code 
    management and project versioning.

    \item \textbf{Development Environment} (Cursor, Claude Code, Jupyter Notebook) --- 
    used for code development, prototyping, and iterative testing 
    of AI pipelines.
\end{itemize}

\subsubsection{New technical skills}

\begin{itemize}
    \item \textbf{Agentic AI Development} --- designing and implementing 
    multi-agent AI systems capable of autonomously planning, retrieving 
    information, reasoning over financial data, and generating structured 
    outputs using LangChain and LangGraph.

    \item \textbf{Retrieval-Augmented Generation (RAG)} --- building RAG 
    pipelines that combine vector-based document retrieval with large 
    language model generation, enabling the agent to ground its qualitative and quantitative
    insights in financial documents.

    \item \textbf{Prompt Engineering} --- designing and optimising prompts 
    for large language models to produce accurate, structured, and 
    contextually relevant fundamental analysis outputs across 
    diverse stock profiles.

    \item \textbf{NLP for Financial Text} --- applying natural language 
    processing techniques (to financial reports, news articles, and 
    industry level reports), entity recognition, 
    and summarisation.

    \item \textbf{Qualitative Research Methods} --- designing and conducting 
    semi-structured interviews with domain experts to evaluate 
    AI-generated qualitative outputs, and thematically analysing 
    interview responses.
\end{itemize}

\subsection{Data requirements}

\begin{itemize}
    \item \textbf{Financial data} (CSE reports / regulatory reports) --- used for retrieving 
    historical financial statements, stock prices, and key financial ratios.

    \item \textbf{Qualitative data} (Reputed news sites, management interviews, macroeconomic reports, industry reviews) 
    --- used for collecting qualitative aspects of the company, its position in the market and future sector forecasts.

    \item \textbf{{Price}} (CSE web) --- used for retrieving 
    current market price.

    \item \textbf{{Research reports}} (Registered Stock Broker reports) --- used for validating computed quantitative values.
\end{itemize}

\subsection{Risks and mitigation}
The following table summarises the major risks identified for the project, together with their corresponding severity, frequency, and mitigation strategies. Severity (S) and Likelihood (L) are evaluated using a scale from 1 to 5, where higher values indicate greater impact or occurrence likelihood.

\begin{longtable}{|p{3.5cm}|p{0.5cm}|p{0.5cm}|p{10cm}|}
    \hline
    \textbf{Risk} & \textbf{S} & \textbf{L} & \textbf{Mitigation Plan} \\
    \hline
    \endfirsthead

    \hline
    \textbf{Risk} & \textbf{S} & \textbf{L} & \textbf{Mitigation Plan} \\
    \hline
    \endhead

    \hline
    \multicolumn{4}{|r|}{\textit{Continued on next page\ldots}} \\
    \hline
    \endfoot

    \hline
    \endlastfoot

    Qualitative Data Quality Risk & 4 & 3 & 
    Implement a Retrieval-Augmented Generation (RAG) pipeline to 
    ground qualitative outputs in verified source documents such 
    as annual reports, industry-wide reports, and broker reports, 
    reducing the likelihood of unsupported or fabricated insights. 
    Outputs will be cross-validated against source documents 
    during testing prior to interview evaluation. \\
    \hline
    Timeline Overrun & 3 & 2 & 
    A detailed project plan with clearly defined milestones and 
    deadlines will be maintained throughout the project. An agile 
    development approach will be adopted, prioritising the delivery 
    of a minimum viable product (MVP) early in the timeline, 
    ensuring that core functionalities are completed before optional 
    enhancements are pursued. Regular supervisor check-ins will be 
    scheduled to monitor progress and flag delays early. \\
    \hline
    Interview Recruitment Challenges & 4 & 4 & 
    Recruitment will commence early in the project timeline to 
    allow sufficient lead time. Multiple recruitment channels will 
    be utilised, including professional networks such as LinkedIn, 
    university alumni networks, and personal industry contacts. 
    Online interview formats (e.g., Zoom, Microsoft Teams) will 
    be offered to reduce participation barriers. A concise 
    participant information sheet outlining the time commitment 
    and research purpose will be prepared to encourage 
    participation. \\
    \hline
    Subjectivity in Qualitative Evaluation & 3 & 2 & 
    A structured interview guide with consistent, pre-defined 
    evaluation criteria will be developed to standardise 
    assessments across participants. Thematic analysis will be 
    employed to systematically identify patterns across interview 
    responses, reducing the impact of individual subjectivity. \\
    \hline
    Agentic AI Pipeline Complexity & 3 & 3 & 
    The pipeline will be developed and tested in modular stages, 
    with each component (data retrieval, preprocessing, reasoning, 
    and output generation) validated independently before 
    integration. Robust error handling and logging mechanisms 
    will be implemented at each stage to detect and isolate 
    failures early. \\
    \hline
    Regulatory Concerns & 1 & 1 & 
    The tool will be explicitly positioned and documented as a 
    research artefact intended for academic evaluation purposes 
    only, and not as a licensed financial advisory service. All 
    outputs will carry appropriate disclaimers. Ethical approval 
    will be sought from the supervisor prior to conducting 
    interviews. Informed consent will be obtained from all 
    participants, and all interview data will be anonymised 
    and stored securely. \\
    \hline
    Computational Resource Constraints & 2 & 2 & 
    Cloud-based LLM APIs (e.g., OpenAI, Anthropic) will be used 
    instead of self-hosted models to minimise local infrastructure 
    requirements. API usage will be carefully monitored and 
    optimised through prompt efficiency and caching of repeated 
    queries to control costs. Free academic credits available 
    through platforms such as Google Cloud, AWS, and OpenAI 
    will be utilised where possible to offset resource expenses. \\
    \hline

    \caption{Risk and mitigation plan}
    \label{tab:risk-mitigation}
\end{longtable}

\section{Chapter summary}

This chapter has established the methodological basis for the research. A pragmatist philosophy, abductive approach, and mixed-methods design were selected to accommodate the dual quantitative and qualitative nature of fundamental analysis. The study follows a design science strategy, centred on the iterative development and evaluation of an Agentic AI artefact targeted at publicly listed companies on the Colombo Stock Exchange. An Agile project management framework guides the development lifecycle, with clearly defined deliverables, resource allocations, and risk mitigation strategies in place to support timely and reliable delivery.
