\chapter{Appendix overview}
Use appendices for supplementary material only. Keep all critical arguments, methodology, and results in the main chapters.

\section{Writing sequence checklist}
\begin{enumerate}
    \item Finalize Chapter 1 (problem, gap, research questions, aim, objectives).
    \item Complete Chapter 2 based on Chapter 1 scope and research questions.
    \item Finalize Chapter 3 and Chapter 4 before implementation details.
    \item Develop Chapter 5 and Chapter 6 iteratively as implementation and experiments progress.
    \item Write Chapter 7 after final evaluation outcomes are stable.
\end{enumerate}

\section{Final quality gate checklist}
\begin{itemize}
    \item Verify numbering flow: cover page unnumbered, front matter roman, main chapters arabic, appendix separate.
    \item Ensure every figure and table has a caption and explanatory reference in text.
    \item Ensure each in-text citation is present in the references list and vice versa.
    \item Verify all research questions and objectives are explicitly answered by the end of Chapter 7.
    \item Confirm critical results remain in main chapters and appendices hold supplementary material only.
\end{itemize}

\section{Detailed concept map}
Insert the full-size concept map used in Chapter 2.

\section{Detailed project schedule}
Insert the full Gantt chart or equivalent timeline artifact.

\section{Additional tables and figures}
Insert large supporting tables, extended results, and non-critical supplementary visuals.

\section{Code snippets and technical artifacts}
Insert selected supplementary code snippets and implementation artifacts if required.
